---
title: "watersword"
format: html
editor: visual
---

## Quarto

```{r library}
library(tidyverse)
```

```{r load-data}
population_projection <- read_csv("data/population_projection.csv")
```


##Executive Summary

In this project we investigate Australia's projected population growth at the national and state levels. Using official population projection datasets and demographic assumptions, we examine the influence of fertility, mortality, and migration on future trends. A linear modelling approach is applied to quantify and compare projection scenarios. Our analysis supports a nuanced understanding of demographic shifts under varying assumptions.

This project investigates Australia's projected population growth, exploring both national and state-specific trends. Our analysis used official projection datasets and demographic assumptions to pinpoint the influence of fertility, mortality, and migration. Through linear modeling, we quantified and contrasted different future scenarios, ultimately providing a nuanced understanding of potential demographic shifts.